---
title: "Spatial Prediction and Model Selection"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
library(ggplot2)
library(readr)
library(scoringRules)
```


## Conditional Multivariate Normal Theory
The conditional distribution, $p(\boldsymbol{Y_1}|
\boldsymbol{Y_2}, \boldsymbol{\beta}, \sigma^2, \phi, \tau^2)$ is normal with:

\vfill


\vfill


\vfill

## Posterior Predictive Distribution

The (posterior) predictive distribution $p(Y(\boldsymbol{s_0})|y)$ can be written as
$$p(Y(\boldsymbol{s_0})|y) = \int p(Y(\boldsymbol{s_0})|y, \boldsymbol{\theta})p( \boldsymbol{\theta}|y) d  \boldsymbol{\theta}$$
where 

\vfill

The posterior predictive distribution 

\vfill


\vfill

## Prediction
We will use cross-validation or a test/training approach to compare predictive models.

\vfill

Consider three data structures: continuous, count, and binary; how should we evaluate predictions in these situations?

\newpage

## Loss Functions
Loss functions penalize predictions that deviate from the true values.

\vfill

For continuous or count data, squared error loss and absolute error loss are common.

\vfill

With binary data, a zero-one loss is frequently used.

\vfill

\vfill

If we think about outcomes distributionally

\vfill


\vfill

The Continuous Rank Probability Score (CRPS) defined as 
$$CRPS(F,y) = \int_{-\infty}^{\infty}\left(F(u) - 1(u \geq y) \right)^2 du,$$
where $F$ is the CDF of the predictive distribution, is a metric that measures distance between an observed value and a distribution.

\newpage

```{r, echo = T}
crps_sample(y = 0, dat = 0)
crps_sample(y = 0, dat = 2)
crps_sample(y = 0, dat = c(2,-2))
crps_sample(y = 0, dat = c(2,0,-2))
crps_sample(y = 0, dat = c(2,0,0,0,0,-2))
```


## CRPS
Consider four situations and sketch the predictive distribution and the resultant CRPS for each scenario. How does the MSE function in each setting?

1. Narrow predictive interval centered around outcome.
\vfill

2. Wide predictive interval centered around outcome.

\vfill

3. Narrow predictive interval with outcome in tail.

\vfill

4. Wide predictive interval with outcome in tail.

\vfill

\newpage

\vfill
```{r}
y <- 0
sample <- rnorm(10000)
f1 <- tibble(data = sample) %>% ggplot(aes(x = data)) + 
  geom_histogram(color = 'grey', fill = 'grey', bins = 100) +
  theme_bw() + geom_vline(xintercept = 0) +
  annotate('text', label = paste("CRPS = ",round(crps_sample(y = y, dat = sample),2)), x = 0, y = 100) + xlim(-10,10)+ ylim(NA,1000)
```


```{r}
y <- 0
sample <- rnorm(10000, sd = 3)
f2 <- tibble(data = sample) %>% ggplot(aes(x = data)) + 
  geom_histogram(color = 'grey', fill = 'grey', bins = 100) +
  theme_bw() + geom_vline(xintercept = 0) +
  annotate('text', label = paste("CRPS = ",round(crps_sample(y = y, dat = sample),2)), x = 0, y = 100) + xlim(-10,10)+ ylim(NA,1000)
```
\vfill


```{r}
y <- 0
sample <- rnorm(10000, mean = -2)
f3 <- tibble(data = sample) %>% ggplot(aes(x = data)) + 
  geom_histogram(color = 'grey', fill = 'grey', bins = 100) +
  theme_bw() + geom_vline(xintercept = 0) +
  annotate('text', label = paste("CRPS = ",round(crps_sample(y = y, dat = sample),2)), x = 0, y = 100) + xlim(-10,10)+ ylim(NA,1000)
```


\vfill



```{r}
y <- 0
sample <- rnorm(10000, mean = -2, sd = 3)
f4 <- tibble(data = sample) %>% ggplot(aes(x = data)) + 
  geom_histogram(color = 'grey', fill = 'grey', bins = 100) +
  theme_bw() + geom_vline(xintercept = 0) +
  annotate('text', label = paste("CRPS = ",round(crps_sample(y = y, dat = sample),2)), x = 0, y = 100) + xlim(-10,10) + ylim(NA,1000)
```

\vfill

```{r}
gridExtra::grid.arrange(f1, f2, f3, f4)
```

